\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-11-17 exps.tex
table
Updated paper

\begin{table}[t]
 \centering
  \twocolumn[
\begin{@twocolumnfalse}
\caption{Benchmark programs using \texttt{delay} constructs} \label{fig:comparison}
\scalebox{0.85}{
\begin{tabular}{| c | c | c | c | c | c | c |}
	\cline{3-7}
	\multicolumn{2}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
								 \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{1-7}
								 \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{1-7}
								 \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{2-7}
											   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{1-7}
								 \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{2-7}
											   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{1-7}
								 \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{2-7}
											   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{1-7}
								 \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{2-7}
											   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{1-7}
								 \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{2-7}
											   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}
% 								& \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8}
% 								&			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8}
% 								&			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8}
% 								&			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8}
% 								&			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8}
% 								&			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
% 	\hline
		\end{tabular}
		}
  \end{@twocolumnfalse}
  ]
\end{table}

@@

We have carried out a number of experiments on a set of applications
with real-time constraints. The benchmark set is shown in
Table~\ref{fig:comparison}. HRCTS is the human response system
described in Figure~\ref{fig:1}. AECS (Access and Environmental Control
System) is an application that controls an intelligent room environment
described in~\cite{aecs_ispa}. Motor is a stepper motor controller
from~\cite{Bourke2009a}. Robot is an industrial automation application
that consists of two manufacturing belts and a robotic arm that splits
goods according to their volumes. HRCTS consists of 2 CDs, but only one
with real-time delays. AECS also has 2 CDs; CD1 consist of a single
delay statement, while CD2 has 2 delay statements. Robot and Motor both
are synchronous programs with 1 and 5 delay statements,
respectively. All these examples are written in the SystemJ
language. Finally, except for HRCTS all other examples have exact (and
deterministic) delay requirements. These delay requirements are shown in
Table~\ref{fig:comparison} (\texttt{M..N}).

@@

We use the \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}, to carry out all our experiments. JOP
implements the JVM specification in hardware directly, and has been
shown to be real-time amenable. We statically estimate the WCRT and BCRT
of these benchmark applications using the real-time analysis tools
provided by the JOP design framework. We disable the \textit{Garbage
  Collector} (GC) in all experiments to avoid unwanted side-effects such
as lengthening clock-domain logical tick times.

As one can see in Table \ref{fig:comparison}, BCRT and WCRT were
calculated for all examples. Given the real-time \mbox{\texttt{delay
    (M..N)}} provided by programmers, required logical delay \texttt{d}
was determined by applying Algorithm~\ref{alg:1}. In HRTCS, real-time
delay of the range 50.3..200.3 (ms) is only satisfied when \emph{d} is
between 1506 and 1787 (as shown previously in
Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
then compiler's job to statically choose one of the values between this
range.

The upper bound \texttt{N} needed to be relaxed for all examples other
than HRCTS to obtain a feasible \texttt{d}. Therefore, Algorithm
\ref{alg:2} is applied to those examples to determine the minimum upper
real-time bound \emph{N} such that there exist a solution for \emph{D}
in Algorithm \ref{alg:1}.  Unlike in HRTCS, however, this algorithm
gives a unique number for logical tick delay \emph{d} such that \emph{d}
= Min(\emph{$S_2$}) = Max(\emph{$S_1$}).  The compiler, therefore, will
choose \emph{d} based on the newly computed \emph{N} for the examples:
Robot, Motor and AECS.

>>>>

\begin{table}[t]
\twocolumn[
\centering
\begin{@twocolumnfalse}
\caption{Benchmark programs using \texttt{delay} constructs} \label{fig:comparison}
\resizebox{18cm}{2.4cm}{
\begin{tabular}{| c | c | c | c | c | c | c |}
	\cline{3-7}
	\multicolumn{2}{c|}{}											& HRTCS 		& Motor 		  & Robot 			  & AECS/CD1  		   & AECS/CD2\\ \hline
								 \multicolumn{2}{|c|}{BCRT} 		&0.0334 ms 		& 0.0361 ms		  & 0.0325 ms		  & 0.1683 ms		   & 0.1881 ms\\ \cline{1-7}
								 \multicolumn{2}{|c|}{WCRT} 		&0.1120 ms 		& 0.1116 ms		  & 0.0601 ms		  & 0.7147 ms		   & 1.0115 ms\\ \cline{1-7}
								 \multirow{2}{*}{Delay 1}	 & M..N &50.3-200.3 ms 	& 2.4-7.4772 ms	  & 1230-2274.0263 ms & 10000-42473.2197 ms& 10000-53772.0045 ms \\ \cline{2-7}
											   & d			 		&1506-1787  	& 67 			  & 37861			  & 59427		       &53160\\ \cline{1-7}
								 \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-5.2452 ms & N/A				  & N/A 			   &10000-53772.0045 ms\\ \cline{2-7}
											   & d			 		& 	N/A	 		& 47			  & N/A				  & N/A       		   &53160\\ \cline{1-7}
								 \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.2232 ms  & N/A				  & N/A       		   &N/A\\ \cline{2-7}
											   & d			 		& 	N/A	 		& 2		      	  &	N/A				  & N/A       		   &N/A\\ \cline{1-7}
								 \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.0044 ms   & N/A				  &	N/A       		   &N/A\\ \cline{2-7}
											   & d				 	& 	N/A	 		& 9				  & N/A				  &	N/A      		   &N/A\\ \cline{1-7}
								 \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-2.3436 ms & N/A				  &	N/A      		   &N/A\\ \cline{2-7}
											   & d			 		& 	N/A	 		& 21 			  &	N/A				  & N/A      		   &N/A\\ \hline
% 	\multirow{12}{*}{TP-JOP} 	& \multicolumn{2}{|c|}{BCRT} 		&0.0028 ms 		& 0.0060 ms		  & 0.0011 ms		  & 0.0532 ms		   & 0.0292 ms\\ \cline{2-8}
% 								& \multicolumn{2}{|c|}{WCRT} 		&0.0329 ms 		& 0.0393 ms		  & 0.0361 ms		  & 0.5072 ms		   & 0.8388 ms\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 1}	 & M..N &50.3-600.3 ms 	& 2.4-15.6265 ms  & 1230-42317.8276 ms& 10000-95358.8578 ms& 10000-287757.9834 ms \\ \cline{3-8}
% 								&			   & d			 		&18209-18232 	& 398 			  & 1171429			  & 188015		       &343054\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 2}	 & M..N & 	N/A 		& 1.667-10.8757 ms& N/A				  & N/A 			   &10000-287757.9834 ms\\ \cline{3-8}
% 								&			   & d			 		& 	N/A	 		& 277			  & N/A				  & N/A       		   &343054\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 3}	 & M..N & 	N/A	 		& 0.05-0.3534 ms  & N/A				  & N/A       		   &N/A\\ \cline{3-8}
% 								&			   & d			 		& 	N/A	 		& 9		      	  &	N/A				  & N/A       		   &N/A\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 4}	 & M..N & 	N/A	 		& 0.3-1.9631 ms   & N/A				  &	N/A       		   &N/A\\ \cline{3-8}
% 								&			   & d				 	& 	N/A	 		& 50			  & N/A				  &	N/A      		   &N/A\\ \cline{2-8}
% 								& \multirow{2}{*}{Delay 5}	 & M..N & 	N/A	 		& 0.733-4.79 ms   & N/A				  &	N/A      		   &N/A\\ \cline{3-8}
% 								&			   & d			 		& 	N/A	 		& 122 			  &	N/A				  & N/A      		   &N/A\\ \cline{2-8}
% 	\hline
\end{tabular}
}
\end{@twocolumnfalse}
]
\end{table}

@@

We have carried out a number of experiments on a set of applications
with real-time constraints. The benchmark set is shown in
Table~\ref{fig:comparison}. HRCTS is the human response system described
in Figure~\ref{fig:1}. AECS (Access and Environmental Control System) is
an application that controls an intelligent room environment described
in~\cite{aecs_ispa}. Motor is a stepper motor controller
from~\cite{Bourke2009a}. Robot is an industrial automation application
that consists of two manufacturing belts and a robotic arm that splits
goods according to their volumes. HRCTS consists of 2 CDs, but only one
with real-time delays. AECS also has 2 CDs; CD1 consist of a single
delay statement, while CD2 has 2 delay statements. Robot and Motor both
are synchronous programs with 1 and 5 delay statements,
respectively. All these examples are written in the SystemJ
language. The requirement to have deterministic or non-deterministic
delays varies between each benchmark. These delay requirements are shown
in Table~\ref{fig:comparison} (\texttt{M..N}).

@@

We use the \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}, to carry out all our experiments. JOP
implements the JVM specification in hardware directly, and has been
shown to be real-time amenable. We statically estimate the WCRT and BCRT
of these benchmark applications using the real-time analysis tools
provided by the JOP design framework. The \textit{Safety Critical Java}
(SCJ) specification we use for real-time analysis avoids garbage
collection (GC) and hence, GC overheads and unpredictability are
completely avoided.

% We disable the \textit{Garbage Collector} (GC) in all experiments to
% avoid unwanted side-effects such as lengthening clock-domain logical
% tick times.

As one can see in Table \ref{fig:comparison}, BCRT and WCRT were
calculated for all examples. Given the real-time \mbox{\texttt{delay
    (M..N)}} provided by programmers, required logical delay \texttt{d}
was determined by applying Algorithm~\ref{alg:1}. In HRTCS, real-time
delay of the range 50.3..200.3 (ms) is only satisfied when \emph{d} is
between 1506 and 1787 (as shown previously in
Section~\ref{sec:intr-real-time}~\ref{sec:find-logic-delay}).  It is
then compiler's job to statically choose one of the values in this
range.

% The upper bound \texttt{N} needed to be relaxed for all examples other
% than HRCTS to obtain a feasible \texttt{d}. Therefore,
Algorithm \ref{alg:2} is applied to those examples, where relaxation was
required, to determine the minimum upper real-time bound \emph{N} such
that there exist a solution for \emph{D} in Algorithm \ref{alg:1}. This
algorithm gives a unique number for logical tick delay \emph{d} such
that \emph{d} = Min(\emph{$S_2$}) = Max(\emph{$S_1$}).  The compiler,
therefore, will choose \emph{d} based on the newly computed \emph{N} for
the examples: Robot, Motor and AECS.

### 2013-11-17 introduction.tex
Updated paper

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delays model relays on
the use of relative instead of absolute time, i.e., a delay in selected
time units is counted from the currently executing SystemJ statement and
(3) delays allow mixing behaviors with real-time features and others
with only logical time and (4) real-time delay does not affect the model
of logical time in SystemJ program as long as the amount of delay is
within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the remaining part of the paper. Our
contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify a
  single \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range, of real-time values for a delay
  statement, e.g., delay statement in Figure~\ref{fig:1}.
\end{enumerate*}

>>>>

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delay model relies on
the use of relative instead of absolute real-time, i.e., a delay in
selected time units is counted from the currently executing SystemJ
statement (3) delays allow mixing behaviors with real-time features and
others with only logical time and (4) real-time delay does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the remaining part of the paper. Our
contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify an
  exact \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range, of real-time values for a delay
  statement, e.g., delay statement in Figure~\ref{fig:1}.
\end{enumerate*}

### 2013-11-17 related_work.tex
Updated paper

A small number of studies have been performed when it comes to
integrating real-time directly as a programming construct in GALS (and
its subset synchronous) reactive languages. The most prominent works in
this area are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
delay in logical ticks \textit{after} allocation and scheduling, it
gives the compiler developers the chance to optimize for many different
criteria such as computation time, energy, power, etc, without worrying
about or violating the real-time delay specification. Other works such
as Quartz~\cite{glog02} also incorporate delay statements in Esterel
using \texttt{pause} constructs, but again, non-maximal parallelism and
non-deterministic delays are not studied. Moreover, Quartz is targeted
at studying timing properties using model checkers rather than
implementation, like us.

>>>>

A number of studies have been performed when it comes to integrating
real-time directly as a programming construct in GALS (and its subset
synchronous) reactive languages. The most prominent works in this area
are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
delay in logical ticks \textit{after} allocation and scheduling, it
gives the compiler developers the chance to optimize for many different
criteria such as computation time, energy, power, etc, without worrying
about or violating the real-time delay specification.

Other works such as Quartz~\cite{glog02} and Bertin
et.al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
statements in Esterel using \texttt{pause} and special \textit{pragmas}
(annotations), respectively.  The Quartz solution is targeted at
studying timing properties using model checkers rather than
implementation, like us. The solution introduced by Bertin
et.al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time
pragmas into Timed automata~\cite{alur94} for model-checking real-time
deadlines. Our solution on the other hand removes all continuous time
elements and produces single discrete time model thereby making the
solution amenable to existing functional and real-time verification
tools developed for reactive languages and more in the spirit of logical
time. Finally, a major difference between Bertin et.al and our approach
is that their approach requires programmers to annotate the time taken
by 'C' code fragments manually, this is extremely error prone, we use
static low-level program analysis to find out such times automatically.
